---
title: "Quantitation of sliding windows and DB analysis"
output: html_notebook

---

### CODE SUMMARY
This code uses a sliding window to quatitate bam file reads. 
Peak quantitation is performed via CSAW.
The normalization is performed using 10k window counts via CSAW.
Differential binding is performed using edgeR.


#### Get list of bam files
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(tidyverse)

# Macs2 beds
bam_dir <- "../data/raw/bams/"
macs_bams <- list.files(path = bam_dir, pattern = ".bam$")

bed_dir <- "../data/interim/bed/"
macs_beds <- list.files(path = bed_dir, pattern = ".narrowPeak.gz$")

# Meta sheet
meta <- read.csv("../src/data/meta.csv", stringsAsFactors = F)
meta$bams <- macs_bams
meta$beds <- macs_beds

# Meta used
meta_used <- meta %>% 
    dplyr::filter(batch == 2, !(sample_name %in% c("456-20-7ft", "615-18A")), !(stage == 'normal'))

meta_used <- meta_used %>% 
    arrange(desc(stage), treatment)


```


#### Count reads into sliding widows
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(csaw)

# Get list of bam files to be used for counting
bam.files <- paste0(bam_dir, meta_used$bams)

# Parameter settings
frag.len <- 185 # determined by cross-correlation in the code down below
win.width <- 150 # recommended by manual for histone marks
wind.param <- readParam(pe="none", minq = 20, dedup = F)

# Count the reads into sliding windows
wind.counts <- windowCounts(bam.files, width=win.width, ext=frag.len, param=wind.param)

colnames(wind.counts) <- meta_used$sample_name

# Determine optimal fragment length using cross-correlation
max.delay <- 500
x <- correlateReads(bam.files, max.delay, param=wind.param)
plot(0:max.delay, x, type="l", ylab="CCF", xlab="Delay (bp)")
print(paste0("Optimal fragment size (bp): ", maximizeCcf(x)))

print("Read counting complete...")
```


#### Filtering out the uninteresteing windows 
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(edgeR)
# Use the count size filtering at 5 reads in average mean library size

# Calculate abundances
abundances <- aveLogCPM(asDGEList(wind.counts))

# Keep windows greater than filter setting
keep <- abundances > aveLogCPM(5, lib.size=mean(wind.counts$totals))
print("Windows kept: ")
summary(keep)

filtered.data <- wind.counts[keep,]

# Export filtered windows as bed file
filt.wind <- filtered.data@rowRanges
export.bed(filt.wind, con = "../data/processed/filtered_windows_rb.bed")

```


#### Normalization
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

# Using 10kb binned for actual normalization
binned <- windowCounts(bam.files, bin=TRUE, width=10000, param=wind.param)
print("10kb bin size: ")
filtered.data <- normFactors(binned, se.out=filtered.data)

# Visualize

par(mfrow=c(1, 2), mar=c(5, 4, 2, 1.5))
adj.counts <- cpm(asDGEList(binned), log=TRUE)
normfacs <- filtered.data$norm.factors
for (i in seq_len(length(bam.files)-1)) {
cur.x <- adj.counts[,1]
cur.y <- adj.counts[,1+i]
smoothScatter(x=(cur.x+cur.y)/2+6*log2(10), y=cur.x-cur.y,
xlab="A", ylab="M", main=paste("1 vs", i+1))
all.dist <- diff(log2(normfacs[c(i+1, 1)]))
abline(h=all.dist, col="red")
}



```


#### PCA analysis
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(plot3D)
library(PCAtools)
library(wesanderson)

wind.dge <- asDGEList(filtered.data,
               group = paste(meta_used$stage, meta_used$treatment, sep = "_"))
colnames(wind.dge) <- meta_used$sample_name

wind.dge$logCPM <- log2(cpm(wind.dge) + 1)

color.grp.pal <- wes_palette("Zissou1", 5, "discrete")[c(5,3)]
color.grp.pal <- c(color.grp.pal, "plum")

#
#Prep the PCA data
pca.wind <- PCAtools::pca(wind.dge$logCPM, metadata = data.frame(meta_used, row.names = colnames(wind.dge$logCPM)))
pca.wind$metadata$Group <- factor(wind.dge$samples$group)

sampkey <- color.grp.pal
names(sampkey) <- unique(pca.wind$metadata$Group)[c(2:3,1)]


#Plot rb samples
rb.dge <- wind.dge[,grep("normal", wind.dge$samples$group, invert = T)]
rb.dge$logCPM <- rb.dge$logCPM[,grep("normal", wind.dge$samples$group, invert = T)]
pca.rb <- PCAtools::pca(rb.dge$logCPM, metadata = data.frame(Group = rb.dge$samples$group, row.names = row.names(rb.dge$samples)))

X='PC1'; Y='PC2'
pca3 <- biplot(pca.rb, x = X, y = Y,
       colby = "Group", colkey = sampkey,
       #lab = "replicate", labSize = 2,
       legendPosition = 'right', 
       title = 'Peak Quant PCA - Retinoblastoma Samples',
       subtitle = paste0(X, ' versus ', Y))

X='PC2'; Y='PC3'
pca4 <- biplot(pca.rb, x = X, y = Y,
       colby = "Group", colkey = sampkey,
       #lab = "replicate", labSize = 2,
       legendPosition = 'right', 
       title = 'Peak Quant PCA - Retinoblastoma Samples',
       subtitle = paste0(X, ' versus ', Y))

print(pca3)
print(pca4)

cowplot::ggsave2(filename = "../data/processed/PCA_peak-rb_PC1-2.pdf", plot = pca3,
                 width = 8, height = 6)
cowplot::ggsave2(filename = "../data/processed/PCA_peak-rb_PC2-3.pdf", plot = pca4,
                 width = 8, height = 6)

```


#### DE Analysis - CSAW manual
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(edgeR)
library(rtracklayer)

y <- asDGEList(filtered.data)
y <- y[,which(meta_used$stage == "blastoma")]
colnames(y) <- meta_used$sample_name[which(meta_used$stage == "blastoma")]

stage_treat <- factor(paste(meta_used$stage[which(meta_used$stage == "blastoma")],
                     meta_used$treatment[which(meta_used$stage == "blastoma")], sep = "_"))
stage_treat[8] <- "blastoma_N"

design <- model.matrix(~stage_treat)
colnames(design) <- c("intercept", "stage_treat")

y <- estimateDisp(y, design)
summary(y$trended.dispersion)

fit <- glmQLFit(y, design, robust=TRUE)
summary(fit$var.post)

par(mfrow=c(1,2))
o <- order(y$AveLogCPM)
plot(y$AveLogCPM[o], sqrt(y$trended.dispersion[o]), type="l", lwd=2,
ylim=c(0, 1), xlab=expression("Ave."~Log[2]~"CPM"),
ylab=("Biological coefficient of variation"))
plotQLDisp(fit)

summary(fit$df.prior)

results <- glmQLFTest(fit, contrast=c(0, 1))
head(results$table)


```


#### Merge windows and calculate p=value
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

# Merge windows
merged <- mergeWindows(filtered.data, tol=200L)
summary(width(merged$regions))
length(merged$regions)
export.bed(merged$regions, con = "../data/processed/merged_windows_rb.bed")


# Get FDR and filter for DE significance
tabcom <- combineTests(merged$ids, results$table)
tabcom <- cbind(data.frame(merged$regions), tabcom)
is.sig.region <- tabcom$FDR <= 0.05
summary(is.sig.region)
tabcom.sig <- tabcom[is.sig.region,]

```


#### Annotate merged windows
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(ChIPseeker)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(xlsx)

txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene

promoter <- getPromoters(TxDb=txdb, upstream=1000, downstream=500)


merged.windAnno <- annotatePeak(merged$regions, tssRegion=c(-1000, 500),
                     TxDb=txdb, annoDb="org.Hs.eg.db")

merged.windAnno.sig <- annotatePeak(merged$regions[is.sig.region], tssRegion=c(-1000, 500),
                     TxDb=txdb, annoDb="org.Hs.eg.db")
# merged.windAnno.sig <- subset(merged.windAnno, subset = is.sig.region)

export.bed(object = merged.windAnno.sig@anno, con = "../data/processed/merged_windows_sigDB_rb.bed")

# Export DB results table
merged_sigDB_output <- cbind(tabcom.sig, as.data.frame(merged.windAnno.sig@anno)[,c(6,12:17)])

write.xlsx(x = merged_sigDB_output, file = "../data/processed/DB_output_results.xlsx", sheetName = "Sig_regions", row.names = F)

```

#### QC for merged windows
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
# Plot annotation bargraph
p <- plotAnnoBar(merged.windAnno)
print(p)

pdf("../data/processed/01_Peak_annot_bar_mergedRb.pdf",
    height = 4, width = 6, useDingbats = F)
print(p)
dev.off()


# Plot annotation pie graph
plotAnnoPie(merged.windAnno)

pdf("../data/processed/01_Peak_annot_pie_mergedRb.pdf",
    height = 4, width = 6, useDingbats = F)
plotAnnoPie(merged.windAnno)
dev.off()

# Plot disrtobution of peaks realtive to TSS
p <- plotDistToTSS(merged.windAnno,
              title="Distribution of H3K27Ac-binding windows\nrelative to TSS")
print(p)

pdf("../data/processed/01_Peak_TSSdist_bar_mergedRb.pdf",
    height = 4, width = 6, useDingbats = F)
print(p)
dev.off()


p <- upsetplot(merged.windAnno)
p <- p +
    labs(title = "H3K27ac clustered peaks annotation")

pdf("../data/processed/01_Peak_upsetplot_mergedRb.pdf",
    height = 4, width = 6, useDingbats = F)
print(p)
dev.off()



#----------------------------
# QC DE Regions

# Plot annotation bargraph
p <- plotAnnoBar(merged.windAnno.sig) 
print(p)

pdf("../data/processed/01_Peak_annot_bar_mergedRb_sig.pdf",
    height = 4, width = 6, useDingbats = F)
print(p)
dev.off()


# Plot disrtobution of peaks realtive to TSS
p <- plotDistToTSS(merged.windAnno.sig,
              title="Distribution of DB H3K27Ac-binding windows\nrelative to TSS in Rb samples")
print(p)

pdf("../data/processed/01_Peak_TSSdist_bar_mergedRb_sig.pdf",
    height = 4, width = 6, useDingbats = F)
print(p)
dev.off()

# Plot annotation pie graph
plotAnnoPie(merged.windAnno.sig)

pdf("../data/processed/01_Peak_annot_pie_mergedRb_sig.pdf",
    height = 4, width = 6, useDingbats = F)
plotAnnoPie(merged.windAnno.sig)
dev.off()

p <- upsetplot(merged.windAnno.sig)
p <- p +
    labs(title = "H3K27ac significant DB clustered peaks annotation")

pdf("../data/processed/01_Peak_upsetplot_mergedRb_sig.pdf",
    height = 4, width = 6, useDingbats = F)
print(p)
dev.off()

```


